Pre-processing high frequency animal tracking data

### 2020-09-16 02_filtering_data.Rmd
new figures filter

# to plot
library(ggplot2)
```

@@

```{r add_outlier}
# add 100 outliers
data_copy <- do_add_outliers(data, p_data = 0.01, std_dev = 0.05)
```

## Removing gross positioning errors

@@

```{r remove_outside_bbox, eval=FALSE}
# remove positions outside a bounding box
data_inside_bbox <- data.table::copy(data_copy)

# NB: set remove_inside to FALSE
data_inside_bbox <- atlastools::atl_filter_bounds(data = data_copy,
                                                  x_range = c(0, 1),
                                                  y_range = c(0, 1),
                                                  remove_inside = FALSE)
```

```{r}
ggplot()+
  geom_path(data = data_copy, aes(x, y), col = "red")+
  geom_path(data = data_inside_bbox, aes(x, y))+
  theme_test()
```

@@

```{r example_remove_outliers, eval=FALSE}
# make a copy of the data
data_many_nbs <- copy(data_inside_bbox)

# get speed and turning angle
data_many_nbs[, `:=`(in_speed = atl_get_speed(data_many_nbs,
                                              type = "in"),
                     out_speed = atl_get_speed(data_many_nbs,
                                               type = "out"),
                     angle = atl_turning_angle(data_many_nbs))]

@@

```{r}
# get 90 and 95 percentile of speed and turning angle
sapply(data_many_nbs[, c("in_speed", "angle")], function(z) {
  quantile(z, probs = c(0.9, 0.95), na.rm = TRUE)
})
```

@@

```{r}
# make a copy of the data
data_non_move <- copy(data_many_nbs)

# filter the copy
data_non_move <- atl_filter_covariates(data_non_move,
            filters = c("(in_speed < 0.001 & out_speed < 0.001) | angle < 40"))
```

```{r}
ggplot()+
  geom_path(data = data_inside_bbox,
            aes(x, y),
            colour = "black")+
  geom_path(data = data_non_move,
            aes(x, y),
            colour = "red")
```

@@

```{r}
ggplot()+
  geom_path(data = data_with_reflection,
            aes(x, y))+
  geom_path(data = data_no_reflection,
            aes(x, y),
            colour = "red")
```

>>>>

# to plot
library(ggplot2)
library(patchwork)
```

```{r}
# source helper functions
source("R/helper_functions.R")
```

@@

```{r add_outlier}
# make a dopy
data_copy <- copy(data)

# add normal error
data_copy[, `:=`(x = do_add_error(x, std_dev = 0.01),
                 y = do_add_error(y, std_dev = 0.005))]
# add 100 outliers
data_copy <- do_add_outliers(data_copy, p_data = 0.005, std_dev = 0.05)
```

## Show data with errors

```{r}
# make list of data to plot
data_plot <- list(data, data_copy)

# make plot list
plots <- lapply(data_plot, function(df) {
  ggplot(data = df)+

    geom_path(aes(x, y),
              col = "red", 
              alpha = 1)+
    geom_point(aes(x, y),
              col = "steelblue", 
              alpha = 0.2,
              pch = 16)+
    theme_test()+
    theme(axis.text = element_blank(),
          axis.title = element_blank())
})

# wrap plots
plot_figure <- wrap_plots(plots) +
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(plot_figure, filename = "figures/fig_add_errors.png", dpi = 300,
       width = 185, height = 90, units = "mm")
```

## Filtering by spatial bounds

@@

```{r remove_outside_bbox, eval=FALSE}
# remove positions outside a bounding box
# NB: set remove_inside to FALSE
data_inside_bbox <- atlastools::atl_filter_bounds(data = data_copy,
                                                  x_range = c(0.25, 0.75),
                                                  y_range = c(0.25, 0.75),
                                                  remove_inside = FALSE)

# data outside a box
data_outside_bbox <- atlastools::atl_filter_bounds(data = data_copy,
                                                  x_range = c(-1, 0.2),
                                                  y_range = c(-1, 0.2),
                                                  remove_inside = TRUE)
```

```{r}
# data to plot
data_plot <- list(data_copy, data_inside_bbox, data_outside_bbox)

# make plots
plots <- lapply(data_plot, function(df) {
  ggplot()+

    geom_path(data = df,
              aes(x, y),
              col = "red", 
              alpha = 1, lwd = 0.2)+
    geom_point(data = df,
               aes(x, y),
              col = "steelblue", 
              alpha = 0.2,
              pch = 16)+
    theme_test()+
    coord_cartesian(xlim = c(0, 1),
                    ylim = c(0, 1)) +
    theme(axis.title = element_blank(),
          axis.text = element_blank())
})

# add elements
plots[[1]] <- plots[[1]] + 
  annotate(xmin = 0.25, xmax = 0.75, ymin = 0.25, ymax = 0.75,
              fill = NA, col = "black", lty = 1, geom = "rect")+
  annotate(x = 0.25, y = 0.8, label = "(b)", 
           fontface = "bold", geom = "text", size = 4)

# add to plot
plots[[1]] <- plots[[1]] + 
  annotate(xmin = -0., xmax = 0.2, ymin = -0.1, ymax = 0.2,
              fill = NA, col = "black", lty = 1, geom = "rect")+
  annotate(x = 0, y = 0.25, label = "(c)", 
           fontface = "bold", geom = "text", size = 4)

# wrap plots
plot_figure <- wrap_plots(plots) +
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

ggsave(plot_figure, filename = "figures/fig_filter_bounds.png", 
       width = 185, height = 65, units = "mm")
```

@@

```{r example_remove_outliers, eval=FALSE}
# get speed and turning angle
data_copy[, `:=`(in_speed = atl_get_speed(data_copy,
                                          type = "in"),
                 out_speed = atl_get_speed(data_copy,
                                           type = "out"),
                 angle = atl_turning_angle(data_copy))]

@@

```{r}
# get 90 and 95 percentile of speed and turning angle
sapply(data_copy[, c("in_speed", "angle")], function(z) {
  quantile(z, probs = c(0.9, 0.95), na.rm = TRUE)
})
```

@@

```{r}
# filter the copy by the 95th percentile
data_filtered <- atl_filter_covariates(data_copy,
            filters = c("(in_speed < 0.029 & out_speed < 0.029) | angle < 40"))
```

```{r}
# data plot
data_plot <- list(data_copy, data_filtered)

plots <- lapply(data_plot, function(df) {
  ggplot()+
     geom_path(data = df,
              aes(x, y),
              col = "red", 
              alpha = 1, lwd = 0.2)+
    geom_point(data = df,
               aes(x, y, 
                   col = (in_speed >= 0.029 & out_speed >= 0.029)),
               shape = ifelse((df$in_speed >= 0.029 & df$out_speed >= 0.029), 
                              16, 1),
               show.legend = F)+
    scale_colour_manual(values = c("steelblue", "red"))+
    theme_test()+
    coord_cartesian(xlim = c(0, 1),
                    ylim = c(0, 1)) +
    theme(axis.title = element_blank(),
          axis.text = element_blank())
})

plot_figure <- wrap_plots(plots) +
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

plot_figure
ggsave(plot_figure, filename = "figures/fig_filter_spikes.png", 
       width = 185, height = 90, units = "mm")
```

@@

```{r}
# data to plot
plot_data <- list(data_with_reflection, data_no_reflection)

# get plots
plots <- lapply(plot_data, function(df) {
  ggplot(data = df)+
    geom_path(aes(x, y),
              col = "red")+
    theme_test()+
    theme(axis.text = element_blank(),
          axis.title = element_blank())
})

patchwork::wrap_plots(plots)+ 
  patchwork::plot_annotation(tag_levels = "a",
                             tag_prefix = "(",
                             tag_suffix = ")")
```
